MLH1 Data Report

### 2019-10-08 doc/Main_Report.rmd
debugging and first results for maternal age effects

pp <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
pp

@@

```{r female.age.effect, echo=FALSE}
#female age effect

@@

```{r mat.age.effect, echo=FALSE}
#compile data from female (and male?) mice with maternal ages.

#start making a general linear model...? for mouse averages

@@

@@

3 different rooms used to house mice. This will be tested in X manner.

>>>>

age.hist <- ggplot(MLH1.imprv[MLH1.imprv$sex == "male", ], aes(age.weeks))+geom_histogram()+facet_wrap(~category)+ggtitle("Male Age and MLH1 counts by subspecies")
age.hist

@@

```{r female.age.effect, echo=FALSE}
#start thinking of the maternal age effect

#female age effect

@@

```{r mat.age.effect, echo=FALSE}
meta.data.file <- meta.data.file[!(is.na(meta.data.file$mouse)|meta.data.file$mouse==""),]
#445? -- probably extras at bottom not good

#how many mice have maternal DOB
meta.data.file.w.mat.DOB <- meta.data.file[!(is.na(meta.data.file$maternal_DOB)|meta.data.file$maternal_DOB==""),]
#151

#format this meta.data.file
source("~./MLH1repo/src/CommonFunc_MLH1repo.R")

meta.data.file.w.mat.DOB <- add_mouse(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_category(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_strain(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_sex(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_subsp(meta.data.file.w.mat.DOB)

#check if date format's been applied before tryung the age functions
str(meta.data.file.w.mat.DOB)

#DOB and Mat.DOB need to be date formated, make sure that the dates are consistant in the original file
#the dates in my file are not all standardized... year is somethimes 2 - sometimes 4
meta.data.file.w.mat.DOB$DOB <- as.Date(meta.data.file.w.mat.DOB$DOB, format= "%m/%d/%Y")#%Y 2000, %y '18,
meta.data.file.w.mat.DOB$maternal_DOB <-as.Date(meta.data.file.w.mat.DOB$maternal_DOB, format= "%m/%d/%Y")

str(meta.data.file.w.mat.DOB)

meta.data.file.w.mat.DOB <- add_euth_date(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_age(meta.data.file.w.mat.DOB)
meta.data.file.w.mat.DOB <- add_mat_age(meta.data.file.w.mat.DOB)

#merge meta data with MLH1 
see.whatcha.do <- merge(AP_mouse_table_w.Ages, meta.data.file.w.mat.DOB, by.x = 'mouse', by.y = 'mouse',all = FALSE )
#remove anything without mean_CO
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mean_co)|see.whatcha.do$mean_co==""),]
see.whatcha.do <- see.whatcha.do[!(is.na(see.whatcha.do$mat.age.days)|see.whatcha.do$mat.age.days==""),]

#visualize the data

meta.data.file.w.mat.DOB.female <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "female",]
meta.data.file.w.mat.DOB.male <- meta.data.file.w.mat.DOB[meta.data.file.w.mat.DOB$sex == "male",]

#hist of data
mat.age.hist.female <- ggplot(meta.data.file.w.mat.DOB.female, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of females")
mat.age.hist.female

mat.age.hist.male <- ggplot(meta.data.file.w.mat.DOB.male, aes(mat.age.days))+geom_histogram()+facet_wrap(~category)+ggtitle("Maternal ages of males")
mat.age.hist.male

#remove mice that might have low quality
see.whatcha.do.HQ <- see.whatcha.do[see.whatcha.do$Ncells > 14,]

#age by mean_co
mat.age.effect.plot.female <- ggplot(see.whatcha.do[see.whatcha.do$sex == "female",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Female Materinal Age and Mean CO (offsping)")+facet_wrap(~subsp.y)
mat.age.effect.plot.female

mat.age.effect.plot.male <- ggplot(see.whatcha.do.HQ[see.whatcha.do.HQ$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")+facet_wrap(~subsp.y)

#
ALLmat.age.effect.plot.male <- ggplot(see.whatcha.do[see.whatcha.do$sex == "male",], aes(y=mean_co, x=mat.age.days, color=(strain.y) ))+geom_point()+ ylim(c(19,34))+
  #trying this
  scale_color_manual(values=colors_of_strains)+ #this list in setupdata
  geom_smooth(method='lm',alpha=0.1, size=0, span=0.5 )+ggtitle("Maternal Age and mean CO, male offspring")

+facet_wrap(~subsp.y)

#is there a correlation with proband age and maternal age?
#there is a big lack of maternal age data...
#some of the values reference the original breeding pair..
#from notes on original BPs(males), assigning maternal ages

@@

There seems to be a negative correlation with

@@

3 different rooms used to house mice. Access softmouse to be fixed. I should have some data on when the switch was. This would be limited to, WSB, LEW and PWD
